### 2012-03-30 paper.tex
Added link to AllRefs.bib to ignore file

\frenchspacing \title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Helle Bunzel, Marjorie Flavin,
    Nir Jaimovich, Lutz Kilian, Michael McCracken, Seth Pruitt, Ross
    Starr, Yixiao Sun, Allan Timmermann, participants at the 2010
    Midwest Economics Association Annual Meetings, the 2010 Joint
    Statistical Meetings, and in many seminars at \textsc{ucsd}, and
    espeically Graham Elliott for their valuable suggestions, feedback
    and advice in writing this paper.  I would also like to thank Amit
    Goyal for providing computer code and data for his 2008 RFS paper
    with Ivo Welch \citep{goyal-welch-2008-rfs}.} \\ Iowa State
  University, Economics Department}
\date{October 24, 2011}

@@

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---when the number of its regressors
increases with the number of observations so that their ratio remains
positive.  We focus on linear regression models estimated with a fixed
training window, but our basic conclusions should be true for other
estimation strategies and models as well.  Under this asymptotic
theory, the \ols\ coefficient estimator is no longer consistent or
asymptotically normal \citep{huber_robust_1973} with positive variance
in the limit.  Even so, we show that the usual \oos\ average is
asymptotically normal and can consistently estimate the difference
between the models' expected loss in periods $T+1,T+2,\dots$
conditional on the on the data available in period $T$, a time-series
analogue to the generalization error.\footnote{See, for example,
  \citet{elemstatlearning}.}  Under these asymptotics, the
generalization error does not converge to the expected performance of
the pseudo-true models, so the in-sample and \oos\ comparisons measure
different quantities and should be expected to give different results
for reasons beyond simple size and power comparisons.  Under our
theory, the model that is closer to the true \dgp\ in population can
forecast worse.  In such a situation, an in-sample comparison would
correctly reject the null hypothesis that the benchmark is true, and
an \oos\ comparison would correctly fail to reject the null that the
benchmark is more accurate.\footnote{In a pair of papers similar to
  ours, \citet{RePEc:fip:fedkrw:rwp09-10,RePEc:fip:fedkrw:rwp09-11}
  study in-sample and \oos\ tests that the larger model has nonzero
  coefficients that are too close to zero to expect the model to
  forecast more accurately.  Like this paper, they argue that the
  larger model can be true but less accurate.  However, they focus on
  an aspect of the \dgp\ that makes this phenomenon likely, while we
  focus on the coefficient estimates that produce less accurate
  forecasts.}

@@

Since the models' coefficients are imprecisely estimated in the limit,
the test sample must be small to ensure that the model estimated over
the training sample is similar to the one that will be estimated over
the full sample.  In particular, $P/T \to 0$ is required for
consistent estimation, and $P^2/T \to 0$ is required for valid
confidence intervals and inference.  For larger $P$, the \oos\
comparisons remain asymptotically normal, but are centered on the
forecasting performance associated with the period-$R$ estimates.  In
practice, researchers typically use large values of $P$, so these
studies may be too pessimistic about their models' future accuracy if
they use the \dmw\ test.  Section~\ref{sec:oostheory} lays out the
asymptotic behavior of the \dmw\ test under this limit theory.

Since the \ols\ coefficient estimator is no longer consistent or
asymptotically normal, the F-test and Wald tests are no longer
asymptotically valid.  Moreover,
\citet{efron_biased_1986,efron_estimation_2004} shows that naive
comparisons of in-sample loss overestimate models' accuracy by a
factor proportional to $K/T$, $K$ being the number of regressors.  But
modifications of the F-test are valid under this asymptotic theory, as
shown by \citet{boos_anova_1995}, \citet{akritas_asymptotics_2000},
\citet{akritas_heteroscedastic_2004}, and
\citet{calhoun_hypothesis_2009}, among others.  Such results establish
that hypotheses about the parameters of the \dgp\ can be tested
in-sample even if one or both models are overfit.  For that very
reason, these tests can not reliably choose a model for forecasting
and will reject the benchmark too often.  We show this analytically in
Section~\ref{sec:insample}.  Moreover, under this asymptotic theory,
many recent \oos\ test statistics, such as those derived by
\cite{clark_tests_2001,clark_evaluating_2005},
\citet{mccracken_asymptotics_2007}, and
\citet{clark_using_2006,clark_approximately_2007} should behave like
the F-test.  These tests are also designed to reject the benchmark
when the alternative model is true, and so they may reject too often
when the benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
\infty$ (with the number of regressors fixed), these statistics behave
like the \dmw\ test when $P$ is small, but should over-reject the
benchmark when $P$ is large.  Section~\ref{sec:mc} presents
simulations that support these claims.

@@

The first assumption is a standard condition on the dependence of the
underlying stochastic array.  We assume that the array is absolutely
regular instead of strong mixing, which is more common in
econometrics, because absolute regularity admits a particular coupling
argument \citep{berbee_random_1979} that is unavailable for strong
mixing sequences.  Please see, for example,
\citet{davidson_stochastic_1994} for a discussion of these mixing
conditions.

Assumption \ref{asmp-2} rules out uninteresting cases where the
forecast error vanishes.
\begin{asmp}\label{asmp-2}
  The variance of $y_{t+\h}$ given $\mathcal{F}_t$ is uniformly
  positive and finite, and all of the eigenvalues of the covariance
  matrix of $x_t$ are uniformly positive and finite as well.  The
  Euclidean norms of the pseudo-true coefficients, $\theta_1$ and
  $\theta_2$, satisfy $|\theta_1|_2 = O(1)$ and $|\theta_2|_2 = O(1)$.
  Moreover, there exist a $\delta > 0$ and finite $\Delta$ such that
  \begin{equation*}
    \lVert \lambda_{\max}(X_{iS_{T}}'X_{iS_{T}}) \rVert_{3+\delta}
    \leq S_{T}\, \Delta \\
    \lVert
    \lambda_{\max}(X_{iS_{T}}'\ep{iS_{T}}\ep{iS_{T}}'X_{iS_{T}})
    \rVert_{3+\delta} \leq S_{T}\, \Delta \\
    \lVert \lambda_{\max}((X_{iS_{T}}'X_{iS_{T}})^{-1}) \rVert_{3+\delta} \leq \Delta/S_{T} \\
    \lVert
    \lambda_{\max}((X_{iS_{T}}'\ep{iS_{T}}\ep{iS_{T}}'X_{iS_{T}})^{-1})\rVert_{3+\delta}
    \leq \Delta/S_{T}
  \end{equation*}
  for all $S_{T} \geq K_{2T}$ and large enough $T$, with $i =
  1,2$,
  \[ X_{iT} \equiv [x_{i1} \quad \dots \quad x_{i,T-\h}]' \]
  and
  \[ \ep{iT} = (\varepsilon_{i,1+\h}, \dots, \varepsilon_{i,T})'.\]
\end{asmp}

>>>>

\frenchspacing \title{Out-of-Sample Comparisons of Overfit Models}
\author{Gray Calhoun\thanks{email: \texttt{gcalhoun@iastate.edu}. I
    would like to thank Julian Betts, Marjorie Flavin, Nir Jaimovich,
    Lutz Kilian, Michael McCracken, Seth Pruitt, Ross Starr, Yixiao
    Sun, Allan Timmermann, participants at the 2010 Midwest Economics
    Association Annual Meetings, the 2010 Joint Statistical Meetings,
    and in many seminars at \textsc{ucsd}, and espeically Graham
    Elliott for their valuable suggestions, feedback and advice in
    writing this paper.  I would also like to thank Amit Goyal for
    providing computer code and data for his 2008 RFS paper with Ivo
    Welch \citep{goyal-welch-2008-rfs}.} \\ Iowa State University,
  Economics Department}
\date{October 24, 2011}

@@

In this paper, we study the ``overfit'' possibility and leave
``instability'' to future research. This paper uses dimension
asymptotics to study the behavior of \oos\ comparisons when at least
one of the models is overfit---when the number of its regressors
increases with the number of observations so that their ratio remains
positive.  We focus on linear regression models estimated with a fixed
training window, but our basic conclusions should be true for other
estimation strategies and models as well.  We show that the usual
\oos\ average is asymptotically normal and can consistently estimate
the difference between the models' expected loss in periods
$T+1,T+2,\dots$ conditional on the on the data available in period
$T$, a time-series analogue to the generalization error.\footnote{See,
  for example, \citet{elemstatlearning}.}  Under these asymptotics,
the generalization error does not converge to the expected performance
of the pseudo-true models, so the in-sample and \oos\
comparisons measure different quantities and should be expected to
give different results for reasons beyond simple size and power
comparisons.  Under our theory, the model that is closer to the true
\dgp\ in population can forecast worse.  In such a situation, an
in-sample comparison would correctly reject the null hypothesis that
the benchmark is true, and an \oos\ comparison would correctly fail to
reject the null that the benchmark is more accurate.\footnote{In a
  pair of papers similar to ours,
  \citet{RePEc:fip:fedkrw:rwp09-10,RePEc:fip:fedkrw:rwp09-11} study
  in-sample and \oos\ tests that the larger model has nonzero
  coefficients that are too close to zero to expect the model to
  forecast more accurately.  Like this paper, they argue that the
  larger model can be true but less accurate.  However, they focus on
  an aspect of the \dgp\ that makes this phenomenon likely, while we
  focus on the coefficient estimates that produce less accurate
  forecasts.}

@@

However, since the models' coefficients are imprecisely
estimated, the test sample must be small to ensure that the model
estimated over the training sample is similar to the one that will be
estimated over the full sample.  In particular, $P/T \to 0$ is
required for consistent estimation, and $P^2/T \to 0$ is required for
valid confidence intervals and inference.  For larger $P$, the \oos\
comparisons remain asymptotically normal, but are centered on the
forecasting performance associated with the period-$R$ estimates.  In
practice, researchers typically use large values of $P$, so these
studies may be too pessimistic about their models' future accuracy if
they use the \dmw\ test.  Section~\ref{sec:oostheory} lays out the
asymptotic behavior of the \dmw\ test under this limit theory.

Under this asymptotic theory, the \ols\ coefficient estimator is no
longer consistent or asymptotically normal \citep{huber_robust_1973},
so this is a useful way to model overfit and imprecisely estimated
models.  Moreover, \citet{efron_biased_1986,efron_estimation_2004}
shows that in-sample comparisons overestimate models' accuracy by a
factor proportional to this ratio.  Despite the non-normality of the
coefficient estimators, modifications of the F-test are valid under
this asymptotic theory, as shown by \citet{boos_anova_1995},
\citet{akritas_asymptotics_2000},
\citet{akritas_heteroscedastic_2004}, and
\citet{calhoun_hypothesis_2009}, among others.  The F-test and Wald
tests are no longer asymptotically valid themselves.  Such results
establish that hypotheses about the parameters of the \dgp\ can be
tested in-sample even if one or both models are overfit.  For that
very reason, these tests can not reliably choose a model for
forecasting and will reject the benchmark too often.  We show this
analytically in Section~\ref{sec:insample}.  Moreover, under this
asymptotic theory, many recent \oos\ test statistics, such as those
derived by \cite{clark_tests_2001,clark_evaluating_2005},
\citet{mccracken_asymptotics_2007}, and
\citet{clark_using_2006,clark_approximately_2007} should behave like
the F-test.  These tests are also designed to reject the benchmark
when the alternative model is true, and so they may reject too often
when the benchmark is false but more accurate.  Obviously, since the
distribution of these statistics converges to the normal when $P/T \to
\infty$ (with the number of regressors fixed), these statistics behave
like the \dmw\ test when $P$ is small, but should over-reject the
benchmark when $P$ is large.  Section~\ref{sec:mc} presents
simulations that support these claims.

@@

The first assumption is a standard condition on the dependence of the
underlying stochastic array.  We assume that the array is absolutely
regular instead of strong mixing, which is more common in
econometrics, because absolute regularity admits a particular coupling
argument \citep{berbee_random_1979} that is unavailable for strong
mixing sequences.

Assumption \ref{asmp-2} rules out uninteresting cases where the
forecast error vanishes.
\begin{asmp}\label{asmp-2}
  The variance of $y_{t+\h}$ given $\mathcal{F}_t$ is uniformly
  positive and finite, and all of the eigenvalues of the covariance
  matrix of $x_t$ are uniformly positive and finite as well.  The
  Euclidean norms of the pseudo-true coefficients, $\theta_1$ and
  $\theta_2$, satisfy $|\theta_1|_2 = O(1)$ and $|\theta_2|_2 = O(1)$.
  Moreover, there exist a $\delta > 0$ and finite $\Delta$ such that
  \begin{equation*}
    \lVert \lambda_{\max}(X_{iS_{T}}'X_{iS_{T}}) \rVert_{4+\delta}
    \leq S_{T}\, \Delta \\
    \lVert
    \lambda_{\max}(X_{iS_{T}}'\ep{iS_{T}}\ep{iS_{T}}'X_{iS_{T}})
    \rVert_{4+\delta} \leq S_{T}\, \Delta \\
    \lVert \lambda_{\max}((X_{iS_{T}}'X_{iS_{T}})^{-1}) \rVert_{4+\delta} \leq \Delta/S_{T} \\
    \lVert
    \lambda_{\max}((X_{iS_{T}}'\ep{iS_{T}}\ep{iS_{T}}'X_{iS_{T}})^{-1})\rVert_{4+\delta}
    \leq \Delta/S_{T}
  \end{equation*}
  for all $S_{T} \geq K_{2T}$ and large enough $T$, with $i =
  1,2$,
  \[ X_{iT} \equiv [x_{i1} \quad \dots \quad x_{i,T-\h}]' \]
  and
  \[ \ep{iT} = (\varepsilon_{i,1+\h}, \dots, \varepsilon_{i,T})'.\]
\end{asmp}
